New cardinality estimation algorithms for HyperLogLog sketches

### 2016-12-21 paper/paper.tex
better description of starting point selection for secant method

Putting all these optimizations together finally gives the new cardinality estimation algorithm presented as \cref{alg:estimate_ml}. The algorithm requires mainly only elementary operations. The square root that appears for the calculation of the stopping limit $\symStopDelta=\symStopEpsilon/\sqrt{\symNumReg}$ can be precomputed, because it only depends on the HyperLogLog precision parameter $\symPrecision$. For large cardinalities it is worth to take the strong \eqref{equ:strong_lower_bound} instead of the weak lower bound \eqref{equ:weak_lower_bound} as initial value, which requires a single logarithm evaluation. The stronger bound is a much better approximation especially for large cardinalities, where the extra logarithm evaluation is amortized by savings in the number of iteration cycles.

>>>>

Putting all these optimizations together finally gives the new cardinality estimation algorithm presented as \cref{alg:estimate_ml}. The algorithm requires mainly only elementary operations. The square root that appears for the calculation of the stopping limit $\symStopDelta=\symStopEpsilon/\sqrt{\symNumReg}$ can be precomputed, because it only depends on the HyperLogLog precision parameter $\symPrecision$.

For very large cardinalities it makes sense to use the strong \eqref{equ:strong_lower_bound} instead of the weak lower bound \eqref{equ:weak_lower_bound} as second starting point for the secant method. The stronger bound is a much better approximation especially for large cardinalities, where the extra logarithm evaluation is amortized by savings in the number of iteration cycles. Therefore, the presented algorithm switches over to the stronger bound, if 
\begin{equation}
\frac{
\sum_{\symRegVal=1}^{\symRegRange}\frac{\symCountVariate_\symRegVal}{2^\symRegVal} + \frac{\symCountVariate_{\symRegRange+1}}{2^\symRegRange}
}{
\sum_{\symRegVal=0}^{\symRegRange}\frac{\symCountVariate_\symRegVal}{2^\symRegVal}
}
> 1.5
\end{equation}
is satisfied. The threshold value of $1.5$ was found to be a reasonable choice in order to reduce the computation time for large cardinalities significantly.
